### 2016-10-20 tex/article.tex
Article++

%% bare_jrnl.tex
%% V1.4b
%% 2015/08/26
%% by Michael Shell
%% see http://www.michaelshell.org/
%% for current contact information.
%%
%% This is a skeleton file demonstrating the use of IEEEtran.cls
%% (requires IEEEtran.cls version 1.8b or later) with an IEEE
%% journal paper.
%%
%% Support sites:
%% http://www.michaelshell.org/tex/ieeetran/
%% http://www.ctan.org/pkg/ieeetran
%% and
%% http://www.ieee.org/

%%*************************************************************************
%% Legal Notice:
%% This code is offered as-is without any warranty either expressed or
%% implied; without even the implied warranty of MERCHANTABILITY or
%% FITNESS FOR A PARTICULAR PURPOSE! 
%% User assumes all risk.
%% In no event shall the IEEE or any contributor to this code be liable for
%% any damages or losses, including, but not limited to, incidental,
%% consequential, or any other damages, resulting from the use or misuse
%% of any information contained here.
%%
%% All comments are the opinions of their respective authors and are not
%% necessarily endorsed by the IEEE.
%%
%% This work is distributed under the LaTeX Project Public License (LPPL)
%% ( http://www.latex-project.org/ ) version 1.3, and may be freely used,
%% distributed and modified. A copy of the LPPL, version 1.3, is included
%% in the base LaTeX documentation of all distributions of LaTeX released
%% 2003/12/01 or later.
%% Retain all contribution notices and credits.
%% ** Modified files should be clearly indicated as such, including  **
%% ** renaming them and changing author support contact information. **
%%*************************************************************************

% *** Authors should verify (and, if needed, correct) their LaTeX system  ***
% *** with the testflow diagnostic prior to trusting their LaTeX platform ***
% *** with production work. The IEEE's font choices and paper sizes can   ***
% *** trigger bugs that do not appear when using other class files.       ***                          ***
% The testflow support page is at:
% http://www.michaelshell.org/tex/testflow/

\documentclass[journal]{IEEEtran}
%
% If IEEEtran.cls has not been installed into the LaTeX system files,
% manually specify the path to it like:
% \documentclass[journal]{../sty/IEEEtran}

% Some very useful LaTeX packages include:
% (uncomment the ones you want to load)

% *** MISC UTILITY PACKAGES ***
%
%\usepackage{ifpdf}
% Heiko Oberdiek's ifpdf.sty is very useful if you need conditional
% compilation based on whether the output is pdf or dvi.
% usage:
% \ifpdf
%   % pdf code
% \else
%   % dvi code
% \fi
% The latest version of ifpdf.sty can be obtained from:
% http://www.ctan.org/pkg/ifpdf
% Also, note that IEEEtran.cls V1.7 and later provides a builtin
% \ifCLASSINFOpdf conditional that works the same way.
% When switching from latex to pdflatex and vice-versa, the compiler may
% have to be run twice to clear warning/error messages.

\usepackage{pbox}

% *** CITATION PACKAGES ***
%
\usepackage{cite}
% cite.sty was written by Donald Arseneau
% V1.6 and later of IEEEtran pre-defines the format of the cite.sty package
% \cite{} output to follow that of the IEEE. Loading the cite package will
% result in citation numbers being automatically sorted and properly
% "compressed/ranged". e.g., [1], [9], [2], [7], [5], [6] without using
% cite.sty will become [1], [2], [5]--[7], [9] using cite.sty. cite.sty's
% \cite will automatically add leading space, if needed. Use cite.sty's
% noadjust option (cite.sty V3.8 and later) if you want to turn this off
% such as if a citation ever needs to be enclosed in parenthesis.
% cite.sty is already installed on most LaTeX systems. Be sure and use
% version 5.0 (2009-03-20) and later if using hyperref.sty.
% The latest version can be obtained at:
% http://www.ctan.org/pkg/cite
% The documentation is contained in the cite.sty file itself.

% *** GRAPHICS RELATED PACKAGES ***
%
\ifCLASSINFOpdf
  \usepackage[pdftex]{graphicx}
  % declare the path(s) where your graphic files are
  \graphicspath{{./figures/}{../jpeg/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  \DeclareGraphicsExtensions{.pdf,.jpeg,.png,.jpg}
\else
  % or other class option (dvipsone, dvipdf, if not using dvips). graphicx
  % will default to the driver specified in the system graphics.cfg if no
  % driver is specified.
   \usepackage[dvips]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../eps/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.eps}
\fi

% graphicx was written by David Carlisle and Sebastian Rahtz. It is
% required if you want graphics, photos, etc. graphicx.sty is already
% installed on most LaTeX systems. The latest version and documentation
% can be obtained at: 
% http://www.ctan.org/pkg/graphicx
% Another good source of documentation is "Using Imported Graphics in
% LaTeX2e" by Keith Reckdahl which can be found at:
% http://www.ctan.org/pkg/epslatex
%
% latex, and pdflatex in dvi mode, support graphics in encapsulated
% postscript (.eps) format. pdflatex in pdf mode supports graphics
% in .pdf, .jpeg, .png and .mps (metapost) formats. Users should ensure
% that all non-photo figures use a vector format (.eps, .pdf, .mps) and
% not a bitmapped formats (.jpeg, .png). The IEEE frowns on bitmapped formats
% which can result in "jaggedy"/blurry rendering of lines and letters as
% well as large increases in file sizes.
%
% You can find documentation about the pdfTeX application at:
% http://www.tug.org/applications/pdftex

% *** MATH PACKAGES ***
%
\usepackage{amsmath}
% A popular package from the American Mathematical Society that provides
% many useful and powerful commands for dealing with mathematics.
%
% Note that the amsmath package sets \interdisplaylinepenalty to 10000
% thus preventing page breaks from occurring within multiline equations. Use:
%\interdisplaylinepenalty=2500
% after loading amsmath to restore such page breaks as IEEEtran.cls normally
% does. amsmath.sty is already installed on most LaTeX systems. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/pkg/amsmath

% *** SPECIALIZED LIST PACKAGES ***
%
%\usepackage{algorithmic}
% algorithmic.sty was written by Peter Williams and Rogerio Brito.
% This package provides an algorithmic environment fo describing algorithms.
% You can use the algorithmic environment in-text or within a figure
% environment to provide for a floating algorithm. Do NOT use the algorithm
% floating environment provided by algorithm.sty (by the same authors) or
% algorithm2e.sty (by Christophe Fiorio) as the IEEE does not use dedicated
% algorithm float types and packages that provide these will not provide
% correct IEEE style captions. The latest version and documentation of
% algorithmic.sty can be obtained at:
% http://www.ctan.org/pkg/algorithms
% Also of interest may be the (relatively newer and more customizable)
% algorithmicx.sty package by Szasz Janos:
% http://www.ctan.org/pkg/algorithmicx

% *** ALIGNMENT PACKAGES ***
%
\usepackage{array}
% Frank Mittelbach's and David Carlisle's array.sty patches and improves
% the standard LaTeX2e array and tabular environments to provide better
% appearance and additional user controls. As the default LaTeX2e table
% generation code is lacking to the point of almost being broken with
% respect to the quality of the end results, all users are strongly
% advised to use an enhanced (at the very least that provided by array.sty)
% set of table tools. array.sty is already installed on most systems. The
% latest version and documentation can be obtained at:
% http://www.ctan.org/pkg/array

% IEEEtran contains the IEEEeqnarray family of commands that can be used to
% generate multiline equations as well as matrices, tables, etc., of high
% quality.

% *** SUBFIGURE PACKAGES ***
%\ifCLASSOPTIONcompsoc
%  \usepackage[caption=false,font=normalsize,labelfont=sf,textfont=sf]{subfig}
%\else
  \usepackage[caption=false,font=footnotesize]{subfig}
%\fi
% subfig.sty, written by Steven Douglas Cochran, is the modern replacement
% for subfigure.sty, the latter of which is no longer maintained and is
% incompatible with some LaTeX packages including fixltx2e. However,
% subfig.sty requires and automatically loads Axel Sommerfeldt's caption.sty
% which will override IEEEtran.cls' handling of captions and this will result
% in non-IEEE style figure/table captions. To prevent this problem, be sure
% and invoke subfig.sty's "caption=false" package option (available since
% subfig.sty version 1.3, 2005/06/28) as this is will preserve IEEEtran.cls
% handling of captions.
% Note that the Computer Society format requires a larger sans serif font
% than the serif footnote size font used in traditional IEEE formatting
% and thus the need to invoke different subfig.sty package options depending
% on whether compsoc mode has been enabled.
%
% The latest version and documentation of subfig.sty can be obtained at:
% http://www.ctan.org/pkg/subfig

% *** FLOAT PACKAGES ***
%
%\usepackage{fixltx2e}
% fixltx2e, the successor to the earlier fix2col.sty, was written by
% Frank Mittelbach and David Carlisle. This package corrects a few problems
% in the LaTeX2e kernel, the most notable of which is that in current
% LaTeX2e releases, the ordering of single and double column floats is not
% guaranteed to be preserved. Thus, an unpatched LaTeX2e can allow a
% single column figure to be placed prior to an earlier double column
% figure.
% Be aware that LaTeX2e kernels dated 2015 and later have fixltx2e.sty's
% corrections already built into the system in which case a warning will
% be issued if an attempt is made to load fixltx2e.sty as it is no longer
% needed.
% The latest version and documentation can be found at:
% http://www.ctan.org/pkg/fixltx2e

\usepackage{stfloats}
% stfloats.sty was written by Sigitas Tolusis. This package gives LaTeX2e
% the ability to do double column floats at the bottom of the page as well
% as the top. (e.g., "\begin{figure*}[!b]" is not normally possible in
% LaTeX2e). It also provides a command:
%\fnbelowfloat
% to enable the placement of footnotes below bottom floats (the standard
% LaTeX2e kernel puts them above bottom floats). This is an invasive package
% which rewrites many portions of the LaTeX2e float routines. It may not work
% with other packages that modify the LaTeX2e float routines. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/pkg/stfloats
% Do not use the stfloats baselinefloat ability as the IEEE does not allow
% \baselineskip to stretch. Authors submitting work to the IEEE should note
% that the IEEE rarely uses double column equations and that authors should try
% to avoid such use. Do not be tempted to use the cuted.sty or midfloat.sty
% packages (also by Sigitas Tolusis) as the IEEE does not format its papers in
% such ways.
% Do not attempt to use stfloats with fixltx2e as they are incompatible.
% Instead, use Morten Hogholm'a dblfloatfix which combines the features
% of both fixltx2e and stfloats:
%
% \usepackage{dblfloatfix}
% The latest version can be found at:
% http://www.ctan.org/pkg/dblfloatfix

%\ifCLASSOPTIONcaptionsoff
%  \usepackage[nomarkers]{endfloat}
% \let\MYoriglatexcaption\caption
% \renewcommand{\caption}[2][\relax]{\MYoriglatexcaption[#2]{#2}}
%\fi
% endfloat.sty was written by James Darrell McCauley, Jeff Goldberg and 
% Axel Sommerfeldt. This package may be useful when used in conjunction with 
% IEEEtran.cls'  captionsoff option. Some IEEE journals/societies require that
% submissions have lists of figures/tables at the end of the paper and that
% figures/tables without any captions are placed on a page by themselves at
% the end of the document. If needed, the draftcls IEEEtran class option or
% \CLASSINPUTbaselinestretch interface can be used to increase the line
% spacing as well. Be sure and use the nomarkers option of endfloat to
% prevent endfloat from "marking" where the figures would have been placed
% in the text. The two hack lines of code above are a slight modification of
% that suggested by in the endfloat docs (section 8.4.1) to ensure that
% the full captions always appear in the list of figures/tables - even if
% the user used the short optional argument of \caption[]{}.
% IEEE papers do not typically make use of \caption[]'s optional argument,
% so this should not be an issue. A similar trick can be used to disable
% captions of packages such as subfig.sty that lack options to turn off
% the subcaptions:
% For subfig.sty:
% \let\MYorigsubfloat\subfloat
% \renewcommand{\subfloat}[2][\relax]{\MYorigsubfloat[]{#2}}
% However, the above trick will not work if both optional arguments of
% the \subfloat command are used. Furthermore, there needs to be a
% description of each subfigure *somewhere* and endfloat does not add
% subfigure captions to its list of figures. Thus, the best approach is to
% avoid the use of subfigure captions (many IEEE journals avoid them anyway)
% and instead reference/explain all the subfigures within the main caption.
% The latest version of endfloat.sty and its documentation can obtained at:
% http://www.ctan.org/pkg/endfloat
%
% The IEEEtran \ifCLASSOPTIONcaptionsoff conditional can also be used
% later in the document, say, to conditionally put the References on a 
% page by themselves.

% *** PDF, URL AND HYPERLINK PACKAGES ***
%
\usepackage{url}
% url.sty was written by Donald Arseneau. It provides better support for
% handling and breaking URLs. url.sty is already installed on most LaTeX
% systems. The latest version and documentation can be obtained at:
% http://www.ctan.org/pkg/url
% Basically, \url{my_url_here}.

% *** Do not adjust lengths that control margins, column widths, etc. ***
% *** Do not use packages that alter fonts (such as pslatex).         ***
% There should be no need to do such things with IEEEtran.cls V1.6 and later.
% (Unless specifically asked to do so by the journal or conference you plan
% to submit to, of course. )

% correct bad hyphenation here
\hyphenation{op-tical net-works semi-conduc-tor}

\begin{document}
%
% paper title
% Titles are generally capitalized except for words such as a, an, and, as,
% at, but, by, for, in, nor, of, on, or, the, to and up, which are usually
% not capitalized unless they are the first or last word of the title.
% Linebreaks \\ can be used within to get better formatting as desired.
% Do not put math or special symbols in the title.
\title{Detection of DCIS and IDC in Whole-Slide\\ H\&E Stained Breast Histopathology Images Using Stacked Convolutional Neural Networks}
%
%
% author names and IEEE memberships
% note positions of commas and nonbreaking spaces ( ~ ) LaTeX will not break
% a structure at a ~ so this keeps an author's name from being broken across
% two lines.
% use \thanks{} to gain access to the first footnote area
% a separate \thanks must be used for each paragraph as LaTeX2e's \thanks
% was not built to handle multiple paragraphs
%

\author{Guido~C.~A.~Zuidhof\textbf{*},
        Babak~Ehteshami~Bejnordi,
        Geert~Litjens,
        and~Jason~Farquhar% <-this % stops a space
\thanks{This paper was written as part of the master's thesis in Artificial Intelligence of G. C. A. Zuidhof. This project was supervised by B. Ehteshami Bejnordi, G. Litjens and J. Farquhar. \textit{Asterisk indicates corresponding author.}}% <-this % stops a space
\thanks{*G. C. A. Zuidhof is a student in the Artificial Intelligence Master's Programme at the Radboud University, 6525HP Nijmegen, The Netherlands (e-mail: guido.zuidhof@student.ru.nl).}% <-this % stops a space
\thanks{B. Ehteshami Bejnordi is with the Diagnostic Image Analysis Group, Radboud University Medical Center, 6500HB Nijmegen, The Netherlands.}%
\thanks{G. Litjens is with the Department of Pathology, Radboud University Medical Center, 6500HB Nijmegen,
The Netherlands.}%
\thanks{J. Farquhar is with the Department of Artificial Intelligence, Radboud University, 6525HP Nijmegen,
The Netherlands and the Donders Institute for Brain, Cognition and Behaviou, 6525EN Nijmegen, The Netherlands.}%

\thanks{First draft, work in progress; October 1, 2016.}}

% note the % following the last \IEEEmembership and also \thanks - 
% these prevent an unwanted space from occurring between the last author name
% and the end of the author line. i.e., if you had this:
% 
% \author{....lastname \thanks{...} \thanks{...} }
%                     ^------------^------------^----Do not want these spaces!
%
% a space would be appended to the last name and could cause every name on that
% line to be shifted left slightly. This is one of those "LaTeX things". For
% instance, "\textbf{A} \textbf{B}" will typeset as "A B" not "AB". To get
% "AB" then you have to do: "\textbf{A}\textbf{B}"
% \thanks is no different in this regard, so shield the last } of each \thanks
% that ends a line with a % and do not let a space in before the next \thanks.
% Spaces after \IEEEmembership other than the last one are OK (and needed) as
% you are supposed to have spaces between the names. For what it is worth,
% this is a minor point as most people would not even notice if the said evil
% space somehow managed to creep in.

% The paper headers
\markboth{IEEE Transactions on Medical Imaging,~Vol.~X, No.~X, October~2016}%
{Zuidhof \MakeLowercase{\textit{et al.}}: Automated detection and characterization of DCIS and IDC in Whole-Slide H\&E Stained Breast Histopathology Images}
% The only time the second header will appear is for the odd numbered pages
% after the title page when using the twoside option.
% 
% *** Note that you probably will NOT want to include the author's ***
% *** name in the headers of peer review papers.                   ***
% You can use \ifCLASSOPTIONpeerreview for conditional compilation here if
% you desire.

% If you want to put a publisher's ID mark on the page you can do it like
% this:
%\IEEEpubid{0000--0000/00\$00.00~\copyright~2015 IEEE}
% Remember, if you use this you must call \IEEEpubidadjcol in the second
% column for its text to clear the IEEEpubid mark.

% use for special paper notices
%\IEEEspecialpapernotice{(Invited Paper)}

% make the title area
\maketitle

% As a general rule, do not put math, special symbols or citations
% in the abstract or keywords.
\begin{abstract}
This paper presents and evaluates a method for detection and characterization of breast malignant lesions in histopathological breast tissue images. The goal of this method is to best classify digitized whole-slide hematoxylin and eosin (H\&E) stained whole-slide images (WSIs) into three classes: benign, ductal carcinoma in situ (DCIS) and invasive ductal carcinoma (IDC). A convolutional neural network (CNN) was trained on small patches of these images, labeled with the class of the center pixel. To distinguish between the DCIS and IDC class the information available in these small patches is often not sufficient. A second CNN was trained on the output of the aforementioned network with a larger patch size to capture a larger context. Sliding this classifier across a whole-slide image (WSI) yields a probability map. From this probability map structural and statistical features were extracted, which were used to train a final classifier to predict the label for the whole-slide image. The system is evaluated on a dataset containing X WSIs of breast tissue using FROC analysis. The results show that the distinction between benign and cancerous WSIs can be predicted with an accuracy of just under 90\%, whereas the three class problem yields an accuracy of 73.4\%.
\end{abstract}

% Note that keywords are not normally used for peerreview papers.
\begin{IEEEkeywords}
Computer-aided diagnosis, DCIS and IDC detection, deep learning, whole-slide imaging.
\end{IEEEkeywords}

% For peer review papers, you can put extra information on the cover
% page as needed:
% \ifCLASSOPTIONpeerreview
% \begin{center} \bfseries EDICS Category: 3-BBND \end{center}
% \fi
%
% For peerreview papers, this IEEEtran command inserts a page break and
% creates the second title. It will be ignored for other modes.
\IEEEpeerreviewmaketitle

%\begin{center} \bfseries \textsc{FIRST DRAFT} \end{center}

\section{Introduction}
% The very first letter is a 2 line initial drop letter followed
% by the rest of the first word in caps.
% 
% form to use if the first word consists of a single letter:
% \IEEEPARstart{A}{demo} file is ....
% 
% form to use if you need the single drop letter followed by
% normal text (unknown if ever used by the IEEE):
% \IEEEPARstart{A}{}demo file is ....
% 
% Some journals put the first two words in caps:
% \IEEEPARstart{T}{his demo} file is ....
% 
% Here we have the typical use of a "T" for an initial drop letter
% and "HIS" in caps to complete the first word.
\IEEEPARstart{B}{reast} cancer is the most common cancer in women~\cite{cancerstats}. Cancers of the breast kill more women than any other form of cancer in all parts of the developing world~\cite{bcstats}. An important tool for the detection and management of breast cancer is analysis of tissue samples under the microscope by a pathologist.

Looking at cancer cells under the microscope, the pathologist searches for certain features that can help predict how likely the cancer is to grow and spread. These features include the spatial arrangement of the cells, morphometric characteristics of the nuclei, whether they form tubules, and how many of the cancer cells are in the process of dividing (mitotic count). These features taken together determine the extent or spread of cancer at the time of diagnosis.

Increasingly, slides of human tissue are digitized instead of being analysed only under a microscope. This has spawned the relatively new field of \emph{digital pathology}. Digital images as opposed to glass slides allow for  easier consultations between pathology experts. An additional advantage of digital images is the opportunity to analyze these images automatically using computer algorithms.

Visual microscopic interpretation of tissue sections is laborious and prone to subjectivity. \emph{Computer-aided diagnosis (CAD)} has a huge potential in alleviating shortcomings of human interpretation and will reduce the workload of the pathologists. As a result, more accurate diagnostic information may be extracted, helping clinicians in selecting the most optimal treatment for individual patients. CAD can facilitate diagnosis by sieving out obviously benign slides and providing quantitative characterization of suspicious areas.

%\textit{TODO: introduce DCIS, IDC a little bit at least.}

\subsection{Goal definition}

We will explore an approach for automatic detection and characterization of these breast malignant lesions in whole-slide images (WSIs) of hematoxylin and eosin (H\&E) stained slides. H\&E staining is the most commonly used staining in medical diagnosis and is often performed as part of standard procedure. It colors cell nuclei blue, and cytoplasm and connective tissue pink. \textit{TODO: say something about other staining/immunohistochemistry methods that can be used, and why not using it may be beneficial (save time, money).}

\begin{figure}[!t]
\centering{}
\includegraphics[width=3.45in]{lesinos_diagrams.png}
\vspace{-0.38cm}\caption{Schematic drawings of mammary ducts. As the cancer progresses malignant and abnormal cells gather within the ducts, the mioepythelium changes and loses its ability to contain these cells within the duct, wihhc renders the cancer invasive. (a) Normal duct. (b) Ductal hyperplasia (benign). (c) Ductal carcinoma in situ (DCIS). (c) Invasive ductal carcinoma (IDC). }
\label{fig_diagrams}
\end{figure}

The final goal of the system is to singly label whole-slide images with one of three classes: \emph{benign, DCIS, or IDC}. Normal tissue and benign lesions are to be assigned to the benign class. A lesion is considered to be a DCIS when there is a proliferation of malignant epithelial cells within the breast parenchymal structures. When these cells are no longer contained within this basement membrane, the carcinoma is invasive and is classified as an invasive ductal carcinoma (IDC) \cite{ellis2010intraductal}. Figure \ref{fig_lesion_examples} shows examples of normal tissue and breast benign, DCIS and IDC lesions.

The method described in this paper to achieve this automatic labeling consists of several building blocks. Figure \ref{fig_overview} provides an overview of the complete system. First, a fully convolutional neural network is used to generate a dense prediction of the whole-slide image. A dense prediction is an assignment of every pixel in an image to one of the labels. To predict the label of a pixel the context (neighborhood area in the image) is important. Due to computational constraints, it is not featsible to use a large patch incorporating large context at high magnifications to train the system. To overcome this limitation we stack a second convolutional neural network on the output of the first, which can capture a larger input area per prediction. From the dense prediction features are extracted, which are then used as feature set for a statistical classifier model to determine the label of the whole-slide image.

To develop and evaluate the system we make use of a large cohort consisting of 221 whole-slide images of H\&E stained slides. We split this dataset into three separate sets, one for fitting classification models, one for intermediate validation and model selection, and one final set for final evaluation of the system; the test set. Table \ref{table_dataset} details the amount of slides in each set. There is no overlap in these sets.

\begin{figure*}[!t]
\centering
\hspace{-0.08in}\includegraphics[width=6in]{system_overview}%
%\label{fig_first_case}}
%\hfil
%\subfloat[Case II]{\includegraphics[width=2.5in]{box}%
\vspace{-0.1cm}\caption{Overview of the proposed system for labeling whole-slide images. (a) Original WSI of breast tissue. (b) Resulting probability map of applying the CNN in a patch-based fashion. (c) Output of the system; a single label for the whole-slide image (either benign, DCIS or IDC).}
\label{fig_overview}
\end{figure*}

\begin{table}[!t]
\renewcommand{\arraystretch}{1.05}
\caption{Dataset details consisting of 221 whole-slide images}
\label{table_dataset}
\centering
\begin{tabular}{|llccc|}
\hline
\textsc{Label}&\textsc{Lesion type}&\hspace{-0.4cm}\textsc{WSI Count}&&\\
\hline
&&\textit{Train}&\hspace{-0.24cm}\textit{Validation}&\textit{Test}\\
\textit{Benign}&&&&\\
&None (normal)& ?&&\\
&Benign lesions& ? &&\\
&\textit{Total}& 50&19&31\\

%\hline
\textit{DCIS}&&&&\\
&Grade I DCIS&  ?&&\\
&Grade II DCIS& ?&&\\
&Grade III DCIS& ?&&\\
&\textit{Total}& 38&11&20\\
%\hline
\textit{IDC}&&&&\\
&IDC& ?&&\\
&\textit{Total}& 30&9&13\\

\textit{\textbf{Total}}&&&&\\
&&\textbf{118}&\textbf{39}&\textbf{64}\\
\hline
\end{tabular}
\end{table}

\begin{figure*}[!t]
\centering
\hspace{0.1cm}\includegraphics[width=7in]{lesion_examples}%
%\label{fig_first_case}}{}
%\hfil
%\subfloat[Case II]{\includegraphics[width=2.5in]{box}%
\caption{Examples from the three classes we aim to distinguish. (a) Normal tissue and benign lesions. From top-to-bottom, left-to-right: ?? with connective tissue surrounding it, mostly fatty tissue, healthy mammal ducts, a ductal hyperplasia which is very similar in appearance to a DCIS lesion. (b) Ductal carcinoma in situ (DCIS), note that the tumor cells are contained within the duct structure. (c) Invasive ductal carcinoma (IDC).}
\label{fig_lesion_examples}
\end{figure*}

\section{Dense prediction}

Convolutional neural networks, as well as other statistical machine learning methods, are bound by computational and memory constraints. The tissue slides are scanned at a high magnification (up to 40X), resulting in very large images. A typical image can be 100,000 by 200,000 pixels, and have a file size of around 20GB uncompressed.

To create a classification for the whole-slide image a patch based method will be employed. A patch is a small sub-image of the whole image. By applying the classifier in a sliding window approach over the original image we can create a prediction map for the whole image.

A trade-off between input size, mini-batch size and model complexity has to be made. If the input size increases, the amount of activations in the feed-forward step of training also increases. These are required for computing the gradient, and thus have to be stored in memory, leading to a larger memory footprint. The amount of images used to determine the error gradient for one update step is called the batch size. A larger batch size generally leads to a less noisy gradient as it is averaged over more data points \cite{bottou2016optimization}. This does not necessarily lead to a faster convergence, or to a better local optimum, as the noise can be beneficial, but in the general sense a larger batch size is beneficial.

\subsection{Incorporating more context}

As a patch is limited in size, it can only contain so much information. One could downsample a larger image to fit a patch, but this would result in the loss of fine-grained detail. To distinguish between benign and cancerous tissue, this fine detail is especially important. To tell apart DCIS from IDC, larger scale architectural features play a more important role.

One approach to incorporate both fine-grained detail as well as larger scale structural information is the use of \emph{multi-scale convolutional neural networks} \cite{multiscale_shen,buyssens2012multiscale}. With this approach an image is offered to one network at different magnifications. Often, there are multiple sub-networks for the different magnification levels. The patches containing a larger context are generally downscalled, which might subsequently lead to a loss of detailed information. The outputs of these sub-networks is then fused, generally using one or more fully connected layers. This allows for training the system as a whole using ordinary backpropagation.

In this paper we explore a different method for incorporating more context, whilst keeping the same level of detail, that of \emph{stacked convolutional neural networks}.

\subsection{Stacked convolutional neural networks}
Convolutional neural networks generally consist of a stack of different types of layers. Some types of layers retain the spatial information of the image, such as convolution layers. In these layers, different filters are convolved over the image. These learned filters are not dependent on the input size and this layer can thus be fed an image of a different size without problems. This characteristic is exploited in \emph{fully convolutional neural networks}, which are the state of the art in segmentation \cite{long2015fully}.

In such a framework, the output segmentation size is only dependent on the input image size.  Therefore, the trained network can be used to segment variable sized images. We will be exploiting this here as well. First, we train a network to classify patches of 224 by 224 pixels. This network will especially learn feature representations that are in the fine detail of the tissue.

Then, we strip this network of its last few layers that do not satisy this fully convolutional requirement. This requirement involves the availability of spatial information and input size independency. We increase the input size of the network to a larger patch size (e.g. 768 by 768 pixels), and do a forward pass through this network. The output is a volume, a feature volume  that resembles an image with many feature maps (or color channels).

We then train another convolutional neural network using this feature volume as its input. The learned weights in the 224 by 224 input-sized network are no longer updated (i.e. they are frozen). The idea is that this stacked network has access to the learned fine detail features, and has very large receptive fields. It can thus learn to exploit both the fine detail, as well as the spatial arrangement of different tissue components.

By training these networks in a tiered manner, we limit the number of trainable parameters to the ones in the stacked network, which will result in a more memory managable training process. This complex network can process patches containg a large receptive field at very high resolution.

\subsection{Architectures}

We will apply two different architectures to the problem of patch classification, with patches of size 224x224 pixels. All implementations of convolutional neural networks were developed using the Theano and Lasagne framework for Python \cite{theano,lasagne}. Full implementations are available on GitHub \footnote{To do: wait for confirmation whether this is allowed.}.

\medskip

The networks we apply for 224x224 patch classification:
\medskip

\subsubsection{Wide Residual Network}
Residual network was the winner of the 2015 ImageNet challenge \cite{resnet}. Other than previous networks that entered this competition, it features no fully connected layer at the end of the network. It was the deepest succesful architecture to date featuring up to 152 layers. It can get away with this depth by using residual learning blocks, where the layers learn residual functions with relation to the input instead of learning unreferenced functions. To the extreme, it makes it easier to learn the identity function if that is optimal for that layer.

In this study we apply an adaptation of this architecture, the wide residual network as proposed by Zagoruyko et al. \cite{wideresnet}. They showed that wider and less deep residual networks outperform their deeper and thinner counterparts both in accuracy and efficiency.

The architecture is a recipe that has three main design choices, the $N$ value that determines the depth, the $k$ value that determines the width (the amount of filters) and the type of ResNet block. Here, we use $N=4$, $k=2$, which was empirically evaluated to fit a decent batch size and train fast. To indicate the architecture with these parameters we will henceforth refer to it as \emph{WRN-4-2}. The ResNet block type is detailed in figure \ref{fig_resnet_block}, this is projection block type B as detailed in the original paper. The resulting architecture is shown in figure \ref{fig_resnet}.
\medskip
\subsubsection{VGG-16}
VGG-16 is a popular architecture for a wide range of image recognition and segmentation problems, partly because of the simplicity in its design \cite{vgg}. It features small 3x3 convolution filters and 2x2 pooling throughout the network and has a depth of 16 layers. The architecture is detailed in figure \ref{fig_vgg}. Unlike the ResNet architecture it does not feature skip connections, also it has a stack of two dense fully connected layers at the end.

\begin{figure*}[!t]
\centering
	\subfloat[{Wide Residual Network architecture. \newline \emph{RB} stands for ResNet Block.}]{\includegraphics[width=2.20in]{architecture_wide_resnet}%
\label{fig_resnet}}
\subfloat[{Architecture of the resnet blocks. \newline The $\otimes$ node indicates an elementwise sum.}]{\hspace{0.8cm}\includegraphics[width=1.8in]{architecture_resnet_block}%
\label{fig_resnet_block}}
\subfloat[VGG-16 model architecture. \newline Dropout with probability of 0.5 was used in the fully connected layers.]{\hspace{0.98cm}\includegraphics[width=2.20in]{architecture_vgg}%
\label{fig_vgg}}
\vspace{-0.12cm}
\caption{Architectures used for 224x224 patch classification. The 1x1 convolution layer in the resnet blocks is only present in the blocks where the input is downsampled, which is every first block with a higher amount of filters in the Wide ResNet architecture (figure \ref{fig_resnet}).}
\label{fig_architectures}
\end{figure*}

\subsection{Preprocessing}
As a preprocessing step, the images are normalized by dividing their pixel value by 255. Then, the mean value of each color channel is subtracted to zero center the data.

\begin{table}[!t]
%% increase table row spacing, adjust to taste
\renewcommand{\arraystretch}{1.1}
% if using array.sty, it might be a good idea to tweak the value of
% \extrarowheight as needed to properly center the text within the cells
\caption{Regions of interest (ROI) per class}
\label{table_roi_count}
\centering
%% Some packages, such as MDW tools, offer better commands for making tables
%% than the plain LaTeX2e tabular which is used here.
\begin{tabular}{|lcc|}
\hline
\textsc{Label}&\textsc{ROI Count}&\\
\hline
&\textit{Train}&\textit{Validation}\\
Benign&5097& 795\\
%\hline
DCIS&1092&  151\\
%\hline
IDC&2307& 432\\
\hline

\end{tabular}
\end{table}

\subsection{Model parameters}
We use a nestov accelerated gradient descent for parameter updates \cite{nesterov}. A starting learning rate of 0.05 was used for the WRN-4-2 model, 0.01 for VGG-16. These values were emperically determined to work well. Rectified linear units were used throughout both networks as nonlinearities. Weights were initialized using He initialization \cite{he_init}. A minibatch size of 22 was used per parameter update step, for determining the performance on the validation set batches of 128 patches were used. Every patch is labeled with the value of the center pixel in the annotations.

The learning rate reduction policy is as follows. The learning rate is multiplied by 0.2 after no better validation accuracy has been observed for 8 epochs, this value is called the \emph{patience}. This patience is increased by 20\% after every reduction in learning rate (rounded up).

To generate one minibatch, first the number of samples per class are determined with uniform probabilities. Then for each class, samples are chosen at random locations from within the annotation of that class. Table \ref{table_roi_count} shows the amount of annotated regions for each of the classes. One epoch generally consists of one pass over the whole training set for parameter updates, followed by a full pass over the validation set to measure convergence. As this sampling can origin from near-infinite amount of pixel locations inside all the annotations for different classes, the concept of epoch is somewhat lost. We will define one epoch to be 300 minibatches for training, and 200 for validation.

Artificially increasing the amount of data by adding variations of the original data can further help train a model that generalizes well. Augmentation consists of applying (random) pertubations to the samples in ways that do not change the label of the samples. It has a regularizing effect which helps prevent overfitting. Especially effective are augmentations that are also realistic examples of real world data.

\medskip

The augmentation methods used are as follows:

\subsubsection{Flips}
The images are randomly mirrored in the X and/or Y direction, both with a 0.5 probability.

\subsubsection{Rotations}
The images are randomly rotated 0, 90, 180, or 270 degrees with equal probability.

\subsubsection{HSV jittering}
HSV is a colorspace that represents a color image in three channels; a \emph{hue} (color), \emph{saturation} (vibrance) and a \emph{value} (brightness) channel. There exists a large variability in the staining of the slides, which shows itself mostly in the hue and saturation channel. We randomly jitter the hue and saturation of an image by multiplying it with a random value between -1.075 and 1.075. The values of all pixels are then clipped between 0 and 1.

\begin{figure}[!t]
\centering{}
\includegraphics[width=3.4in]{augment_examples}
\caption{Examples of augmentations applied to patches of 224 by 224 pixels. In the left column the original image is shown, in the other columns augmented versions of this same image are shown. Note the slight shift in hue and saturation for some of the augmentations.}
\label{fig_augment}
\end{figure}

\medskip

Another commonly used data augmentation method is also zooming. This was not used here, as the size of the nuclei plays a role in determining the class label. Elastic distortions were succesfully applied as part of the data augmentation strategy in other applications of convolutional neural networks \cite{elastransforms}, also in the medical imaging domain (cytology) \cite{unet}. The irregularity of the sizes of the nuclei, as well as the architectural features are important biomarkers that can be used to distinguish between benign and cancerous lesions. Applying these elastic distortions would likely contaminate this information, and as such, it was not used here.

\subsection{Results}

\input{tables/table_accuracy_224}

We trained four different convolutional neural networks on patches of 224 by 224 pixels. Two of each architecture (WRN-4-2 and VGG-16). The results are detailed in table \ref{table_results_224}. The results in this table show the best epoch in regards of validation accuracy. This value is based on the average of 200 batches of 128 randomly sampled patches with a uniform class probability, and may thus be quite noisy.

From these results we can make these observations:

\subsubsection{Architecture performance}
On the two class problem of benign versus cancer patches, the wide resnet outperformed the VGG-16 architecture with an accuracy of 0.9241 as opposed to 0.9237. This difference is, however, so small that it is likely insignificant. On the three class problem however, the opposite is observed. Here the WRN-4-2 network achieved an accuracy of 0.7995, much lower than that of the VGG-16 network which achieved a 0.8245 accuracy. Figure \ref{fig_3class} shows the training and validation accuracy over time. The validation accuracy of the wide resnet is very noisy, likely the learning rate was too high.

\subsubsection{Two class and three class performance}
Accuracy on the three class problem was much lower. This is likely due to the lack of context available in the small 224 by 224 patches which is paramount in distinguishing the DCIS and IDC class. As can be seen in figure \ref{fig_2class} and \ref{fig_3class}, the difference in training and validation set accuracy was much greater for the three class networks. This indicates a larger overfit on the training set, which is to be expected when not enough features (or context) is available to learn generalizable patterns.

Note that these are only intermediate results as these networks will not be used for dense prediction individually. Instead, we train a network on top of the output of these models.

\begin{figure}[!t]
\centering{}
\hspace{-0.3cm}\includegraphics[width=3.5in]{2class_performance}
\vspace{-0.35cm}\caption{Performance of networks trained on the two class problem of benign versus cancerous patches. The dashed line shows the performance on the images from the train set it was shown that epoch.}
\label{fig_2class}
\end{figure}

\begin{figure}[!t]
\centering{}
\hspace{-0.3cm}\includegraphics[width=3.5in]{3class_performance}
\vspace{-0.35cm}\caption{Performance of networks trained on all three classes. The dashed line shows the performance on the images from the train set it was shown that epoch. The validation accuracy of the wide resnet is especially noisy.}
\label{fig_3class}
\end{figure}

\section{Stacked architecture}

In order to increase the context available for dense prediction we stack a second CNN on top of the last fully convolutional layer of the WRN-4-2 224x224 network. We compare the performance between stacking on top of a 2 class network versus a 3 class network. Also, we will look into the effect of patch size on network performance by comparing 512x512, 768x768 and 1024x1024 input-sized networks.

In figure \ref{fig_stacked}, the stacked architecture is shown. It is a hybrid between the wide resnet architecture and the VGG architecture. This architecture is a fully convolutional network, which allows for faster dense prediction due to the possibility of re-using overlapping predictions.

We use a starting learning rate of 0.005 with the same learning rate decay strategy as the network trained on patches of size 224x224. Also, the same augmentations are used. For the 512x512 and 768x768 networks a training and validation minibatch size of 18 and 40 was used, respectively. We define one epoch to be 100 of these update steps, and average over 75 of these validation minibatches.

The network trained on 1024x1024 patches has a higher memory footprint that is too high for these minibatch sizes, instead one minibatch consists of 10 patches. For validation we average over 125 batches of 22 patches.

\begin{figure}[!t]
\centering{}
\includegraphics[width=2.2in]{architecture_stacked}
\vspace{-0.15cm}\caption{Architecture of the stacked 768x768 input-sized network. The weights of the components with the dotted outlines are taken from the previously trained 224x224 patch model, and are no longer updated (these are frozen).}
\label{fig_stacked}
\end{figure}

\subsection{Results}

\input{tables/table_accuracy_stacked}

Figure \ref{fig_stack} shows the performance of the network during training. Again, we observe a very noisy validation accuracy over the course of the training. The network stacked on top of the 2 class network shows a less noisy accuracy over time as opposed to that of the 3 class network. This higher noise for this network may be due to the lack of convergence of the network it is stacked on top of. The network weights at the best epoch were selected, which was epoch 40 (see figure for this network \ref{fig_3class}). The stacked network on top of this three class network did however outperform this network by around 2\% accuracy. The complete set of results are shown in table \ref{table_results_stacked}.

The accuracy of all three networks is much greater than that of the original networks trained on patches of size 224x224 they are stacked on top of. We can also see a clear improvement in accuracy as the patch size is 
increased, hinting towards this increase in accuracy not just being attributable to the increase in network depth. The 1024x1024 outperformed the smaller architectures (0.9135 versus 0.9050 and 0.8723 for the 768x768 and 512x512 network), despite the smaller batch size. This network is however computationally much more demanding to both train and make predictions with. For this reason we select the 768x768 network for dense prediction.

%\begin{figure}[!t]
%\centering{}
%\hspace{-0.3cm}\includegraphics[width=3.5in]{768_performance}
%\vspace{-0.35cm}\caption{Performance of stacked networks, either on the 2 class WRN-4-2 and the 3 class WRN-4-2 model.}
%\label{fig_stack}
%\end{figure}

\begin{figure}[!t]
\centering{}
\hspace{-0.45cm}\includegraphics[width=3.65in]{768_performance_dual}
\vspace{-0.35cm}\caption{Performance of stacked networks. The left graph shows the performance of the 768x768 networks stacked on the 2 class WRN-4-2 and the 3 class WRN-4-2 model. The right graph shows the performance of stacked networks with various input sizes, all stacked on the 3 class WRN-4-2 model. The dotted lines indicate performance on the train set.}
\label{fig_stack}
\end{figure}

%\begin{figure}[!t]
%\centering{}
%\hspace{-0.3cm}\includegraphics[width=3.5in]{patch_sizes_performance}
%\vspace{-0.35cm}\caption{Performance of networks with different input sizes, all stacked on the 3 class WRN-4-2 model.}
%\label{fig_stack_different_input_sizes}
%\end{figure}

\subsection{Qualitative results}
The model we select for dense prediction will not create a perfect probability map. Looking at the confusion matrix in table \ref{table_confusion_768}, we can expect most errors in misclassified patches to be between benign and the DCIS class. As for the DCIS versus IDC distinction, patches are more often misclassified as being DCIS when they are actually IDC, than the other way around. Figure \ref{fig_misclassifications} shows a selection of both correctly and incorrectly classified patches.

\begin{table}[h]
%% increase table row spacing, adjust to taste
\renewcommand{\arraystretch}{1.21}
% if using array.sty, it might be a good idea to tweak the value of
% \extrarowheight as needed to properly center the text within the cells{}
\caption{Confusion matrix of predictions of model used for dense prediction, based on 72,000 random validation set patches }
\label{table_confusion_768}
\centering
%% Some packages, such as MDW tools, offer better commands for making tables
%% than the plain LaTeX2e tabular which is used here.
\begin{tabular}{|l|rrr|}
\hline
\textsc{}&\textsc{Benign}&\textsc{DCIS}&\textsc{IDC}\\
\hline
\textsc{Benign}&0.878&0.088&0.034\\
\textsc{DCIS}&0.068& 0.904&0.028\\
\textsc{IDC}&0.015& 0.081&0.904\\
\hline

\end{tabular}
\end{table}

\begin{figure}[!t]
\vspace{-0.435cm}
\centering{}
\hspace{-0.1cm}\includegraphics[width=3.5in]{figures/misclassifications.png}
\vspace{-0.55cm}\caption{Hand selected examples of patches that were correctly and incorrectly classified. (a) Correct benign. (b) Benign patch misclassified as DCIS. (c) Correct DCIS. (d) DCIS misclassified as benign. (e) Correct IDC. (f) IDC misclassified as DCIS. }
\label{fig_misclassifications}
\end{figure}

\subsection{Dense prediction}
To create the dense prediction, we slide the patch clasifier over the whole-slide image until the full image has been covered. We select the best performing 768x768 stacked CNN for this purpose. In our case this is the model stacked on the three class WRN-4-2 network trained on 224x224 patches.

Whilst this architecture is fully convolutional, we do not exploit this by further increasing the input size. Throughout the network padding is applied, which can potentially lead to learned border effects in the prediction which may not be present in the same way for a larger input size. Whilst this may not necessarily hurt performance, we do not utilize this option to make sure this does not influence results. We move the classifier by 256 pixels (e.g. a stride of 256) until the whole image has been covered to create a probability map with a decent probability.

Figure \ref{fig_wsi_examples} shows examples of dense predictions of WSIs with these parameters. It yields a course segmentation of the different types of tissue.
%\vspace{-1cm}

\begin{figure}[h]
\centering{}
\hspace{0.09cm}\includegraphics[width=3.35in]{wsi_predictions_examples}
\vspace{-0.15cm}\caption{Randomly selected dense prediction maps. The red, green and blue color channels indicate the probability of benign, DCIS and IDC class respectively. Black pixels are fully background patches that were not classified.}
\label{fig_wsi_examples}
\end{figure}

% An example of a floating figure using the graphicx package.
% Note that \label must occur AFTER (or within) \caption.
% For figures, \caption should occur after the \includegraphics.
% Note that IEEEtran v1.7 and later has special internal code that
% is designed to preserve the operation of \label within \caption
% even when the captionsoff option is in effect. However, because
% of issues like this, it may be the safest practice to put all your
% \label just after \caption rather than within \caption{}.
%
% Reminder: the "draftcls" or "draftclsnofoot", not "draft", class
% option should be used if it is desired that the figures are to be
% displayed while in draft mode.
%
%\begin{figure}[!t]
%\centering
%\includegraphics[width=2.5in]{myfigure}
% where an .eps filename suffix will be assumed under latex, 
% and a .pdf suffix will be assumed for pdflatex; or what has been declared
% via \DeclareGraphicsExtensions.
%\caption{Simulation results for the network.}
%\label{fig_sim}
%\end{figure}

% Note that the IEEE typically puts floats only at the top, even when this
% results in a large percentage of a column being occupied by floats.

% An example of a double column floating figure using two subfigures.
% (The subfig.sty package must be loaded for this to work.)
% The subfigure \label commands are set within each subfloat command,
% and the \label for the overall figure must come after \caption.
% \hfil is used as a separator to get equal spacing.
% Watch out that the combined width of all the subfigures on a 
% line do not exceed the text width or a line break will occur.
%
%\begin{figure*}[!t]
%\centering
%\subfloat[Case I]{\includegraphics[width=2.5in]{box}%
%\label{fig_first_case}}
%\hfil
%\subfloat[Case II]{\includegraphics[width=2.5in]{box}%
%\label{fig_second_case}}
%\caption{Simulation results for the network.}
%\label{fig_sim}
%\end{figure*}
%
% Note that often IEEE papers with subfigures do not employ subfigure
% captions (using the optional argument to \subfloat[]), but instead will
% reference/describe all of them (a), (b), etc., within the main caption.
% Be aware that for subfig.sty to generate the (a), (b), etc., subfigure
% labels, the optional argument to \subfloat must be present. If a
% subcaption is not desired, just leave its contents blank,
% e.g., \subfloat[].

% An example of a floating table. Note that, for IEEE style tables, the
% \caption command should come BEFORE the table and, given that table
% captions serve much like titles, are usually capitalized except for words
% such as a, an, and, as, at, but, by, for, in, nor, of, on, or, the, to
% and up, which are usually not capitalized unless they are the first or
% last word of the caption. Table text will default to \footnotesize as
% the IEEE normally uses this smaller font for tables.
% The \label must come after \caption as always.
%
%\begin{table}[!t]
%% increase table row spacing, adjust to taste
%\renewcommand{\arraystretch}{1.3}
% if using array.sty, it might be a good idea to tweak the value of
% \extrarowheight as needed to properly center the text within the cells
%\caption{An Example of a Table}
%\label{table_example}
%\centering
%% Some packages, such as MDW tools, offer better commands for making tables
%% than the plain LaTeX2e tabular which is used here.
%\begin{tabular}{|c||c|}
%\hline
%One & Two\\
%\hline
%Three & Four\\
%\hline
%\end{tabular}
%\end{table}

% Note that the IEEE does not put floats in the very first column
% - or typically anywhere on the first page for that matter. Also,
% in-text middle ("here") positioning is typically not used, but it
% is allowed and encouraged for Computer Society conferences (but
% not Computer Society journals). Most IEEE journals/conferences use
% top floats exclusively. 
% Note that, LaTeX2e, unlike IEEE journals/conferences, places
% footnotes above bottom floats. This can be corrected via the
% \fnbelowfloat command of the stfloats package.

\section{Whole-slide image labeling}
The second goal is to classify the whole-slide image into benign, DCIS or DCIS classes. We create a dense prediction using the stacked convolutional neural network trained previously and extract features from this probability map. These features are then used to train a classifier for classifying the whole-slide image. We treat it as a multiclass problem in which one label is assigned to a whole-slide image. Note that an image may contain both DCIS and IDC lesions. When this is the case, it should receive the IDC label.

The probability map is first thresholded giving every pixel a single label. The label with the highest probability is selected for every pixel. Background pixels are discarded in the calculation of all features. 
We extract both features that describe the image as a whole, as well as those that capture the spatial arrangement of tissue components. One WSI may contain multiple tissue sections, often the next slice of tissue. An example of this can be seen the second prediction in figure \ref{fig_wsi_examples}. Those features that are not full image descriptors are determined independently for each of these tissue sections.

The features we extract are the following:

\subsubsection{Full image descriptors}
The first set of features we extract are those that describe the image as a whole. We determine the fraction of pixels that are benign, DCIS, IDC and cancerous (DCIS and IDC combined).Then, we also determine the fraction of DCIS and IDC w.r.t. all cancerous pixels.

\subsubsection{Voronoi diagram features}
A voronoi diagram is a partitioning of an area of pixels based on the distance to certain points. Conceptually these regions are the result of region growing from certain seed points. To create the seed points we first create a binary image of both the DCIS and IDC class, we then determine the connected components. The center of these components is used as seed for growing (a voronoi diagram is created for both classes indepently and its features are both saved). For each of the voronoi regions we extract the following metrics: The area of the region, the eccentricity, the area of this region divided by the total tissue area and the fraction of this area filled with this label's prediction. Per WSI we determine the mean, median and standard deviation of these voronoi area metrics. Also, we determine the largest single region and add this to the feature set.

Additionaly, we determine the average amount of DCIS and IDC connected components per tissue section.

\subsubsection{Delaunay triangulation features}
For each of the connected components in both the DCIS and IDC masked image we also construct a delaunay triangulation graph using the connected components' center point as nodes. Figure \ref{fig_delaunay} shows an example of such a delaunay triangulation graph overlayed over the dense prediction. For each of the nodes we count the amount of neighbors that are closer than a certain threshold (we used 24 pixels). Also, the average distance of these neighbors is determined for each node. Here we determine the mean, median and standard deviation of these values and use these as features. We also determine the highest average node distance in the whole image.

\subsubsection{Convex hull area features}
IDC lesions often form a single large connected mass, an example of which can be seen in the third and fifth image in figure \ref{fig_wsi_examples}. Manual inspection of the probability maps showed that when these areas are more sparse, they are often false positives. To capture this in a feature we filter the small connected components (less than 12 pixels in size), and then construct a convex hull of the remaining connected components. We then determine the fraction of IDC labeled pixels in the area contained in the convex hull. If multiple tissue sections are present, we average over the different tissue sections which contain IDC.

\begin{figure}[!t]
\centering{}
\includegraphics[width=3.35in]{feature_delaunay}
\vspace{-0.15cm}\caption{Delaunay triangulation of DCIS lesions. Benign tissue is shown in light blue, DCIS annotations are shown in yellow. The small dark spots are pixels labeled with the IDC class. For each of the red nodes the amount of neighbors and the average distance to these is determined.}
\label{fig_delaunay}
\end{figure}

The resulting feature vector contains 54 features. We use these features as input for a random forest classification model as proposed by Breiman \cite{breiman2001random}. We train a forest of 128 random decision trees, each of which receives the full feature set as input. A Scikit-learn implementation of this method was used\cite{scikit-learn}.

\subsection{Evaluation}

We tweak and evaluate the models by applying cross validation on the combined set of train and validation whole slide images. Because of the small amount of images, we adopt a leave-one out cross validation strategy. In such a scheme,  one image is removed from the dataset, a model is trained on all other images, which is then used to predict this single image. This is repeated until a prediction is made for every image.  This is equivalent to K-fold cross validation in which K is equal to the amount of samples.

As a final evaluation, we classify images from the held-out test subset of the data. This data has not been used to train any of the models in the whole pipeline, and will thus be a good measure of the generalisability of the method. Again, see table \ref{table_dataset} for the amount of whole-slide images per class. It is noteworthy that the performance on this combined set of train and validation set images may be inflated as the CNNs were trained on patches from this train set and are thus likely to have better dense predictions of these slides. This effect is magnified by the train set being around 3 times larger in count than the validation set.

We evaluate our models on three metrics:
\subsubsection{Accuracy}
The accuracy is simply the fraction of correct predictions. Because of the class imbalance present in our dataset (see table \ref{table_dataset}) this measure can give a distorted view of the predictive value of the model.

\subsubsection{Cohen's kappa coefficient}
Cohen's kappa is a statistic that is generally used to measure inter-annotator agreement. It is not sensitive to class imbalance. It is defined as

$$\kappa = \frac{p_o-p_e}{1-p_e},$$

where $p_o$ is the observed agreement ratio, and $p_e$ is the expected agreement under random assignment of labels \cite{artstein2008inter}. This measure can range between -1 and 1, where 0 corresponds to chance agreement and higher is better.

\subsubsection{Area under curve (AUC)}
As the discrimination threshold is varied, the true and false positive rate of a model will vary. An operating point can be selected at which a point is assigned to a class or not. We create a receiver operating characteristic (ROC) curve that shows the performance as this threshold is changed. The area under this curve is \textit{``equal to the probability that a classifier will rank a randomly chosen positive instance higher than a randomly chosen negative one'' \cite{Fawcett:2006:IRA:1159473.1159475}.} This provides a good measure of the overall performance of the model and its ability to capture uncertainty in its labeling.

Whilst this can be calculated for the multiclass problem by composing the problem in a one-versus-rest scheme, we choose only to determine this for the binary problem of benign versus cancer WSIs. This decision is made because of the very low amount of samples with DCIS and IDC labels, which would result in a very blocky and inaccurate curve.

\subsection{Results}
\input{tables/table_rf}

Accuracy and the kappa coefficient on the held out test set was much lower than that of the cross validated results for the three class problem. Taking the class imbalance into account, the confusion matrix of the test set predictions in table \ref{table_confusion_test} shows that especially errors between the DCIS and IDC class are common. This explains the large difference between the two and three class model performance metrics.

\begin{table}[h]
\renewcommand{\arraystretch}{1.21}
\caption{Confusion matrix of test set predictions}
\label{table_confusion_test}
\centering
\begin{tabular}{|l|rrr|}
\hline
\textsc{}&\textsc{Benign}&\textsc{DCIS}&\textsc{IDC}\\
\hline
\textsc{Benign}&27&4&0\\
\textsc{DCIS}&4& 11&5\\
\textsc{IDC}&0& 4&9\\
\hline

\end{tabular}
\end{table}

\begin{figure}[!t]
\centering{}
\hspace{-0.3cm}\includegraphics[width=3.5in]{roc}
\vspace{-0.35cm}\caption{Receiver operator characteristic (ROC) curve. The solid and dashed line are for the independent test set and cross validated train set respectively.}
\label{fig_roc}
\end{figure}

\section{Conclusion}

As the dense prediction improves, we can expect the whole-slide image prediction to improve with it. With a (near) perfect dense prediction, predicting the label of the whole-slide image becomes trivial.
The conclusion goes here.

% if have a single appendix:
%\appendix[Proof of the Zonklar Equations]
% or
%\appendix  % for no appendix heading
% do not use \section anymore after \appendix, only \section*
% is possibly needed

% use appendices with more than one appendix
% then use \section to start each appendix
% you must declare a \section before using any
% \subsection or using \label (\appendices by itself
% starts a section numbered zero.)
%

%\appendices
%\section{Proof of the First Zonklar Equation}
%Appendix one text goes here.

% you can choose not to have a title for an appendix
% if you want by leaving the argument blank
%\section{}
%Appendix two text goes here.

% use section* for acknowledgment
%\section*{Acknowledgment}

%The authors would like to thank...

% Can use something like this to put references on a page
% by themselves when using endfloat and the captionsoff option.
\ifCLASSOPTIONcaptionsoff
  \newpage
\fi

% trigger a \newpage just before the given reference
% number - used to balance the columns on the last page
% adjust value as needed - may need to be readjusted if
% the document is modified later
%\IEEEtriggeratref{8}
% The "triggered" command can be changed if desired:
%\IEEEtriggercmd{\enlargethispage{-5in}}

% references section

% can use a bibliography generated by BibTeX as a .bbl file
% BibTeX documentation can be easily obtained at:
% http://mirror.ctan.org/biblio/bibtex/contrib/doc/
% The IEEEtran BibTeX style support page is at:
% http://www.michaelshell.org/tex/ieeetran/bibtex/
\bibliographystyle{IEEEtran}
% argument is your BibTeX string definitions and bibliography database(s)
\bibliography{IEEEabrv,bib/ref}
%,
% <OR> manually copy in the resultant .bbl file
% set second argument of \begin ,to the number of references
% (used to reserve space for the reference number labe,ls box)
%\begin{thebibliography}{1}

%\bibitem{IEEEhowto:kopka}
%H.~Kopka and P.~W. Daly, \emph{A Guide to \LaTeX}, 3rd~ed.\hskip 1em plus
%  0.5em minus 0.4em\relax Harlow, England: Addison-Wesley, 1999.

%\end{thebibliography}

% biography section
% 
% If you have an EPS/PDF photo (graphicx package needed) extra braces are
% needed around the contents of the optional argument to biography to prevent
% the LaTeX parser from getting confused when it sees the complicated
% \includegraphics command within an optional argument. (You could create
% your own custom macro containing the \includegraphics command to make things
% simpler here.)
%\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{mshell}}]{Michael Shell}
% or if you just want to reserve a space for a photo:

%\begin{IEEEbiography}{Michael Shell}
%Biography text here.
%\end{IEEEbiography}

% if you will not have a photo at all:
%\begin{IEEEbiographynophoto}{John Doe}
%Biography text here.
%\end{IEEEbiographynophoto}

% insert where needed to balance the two columns on the last page with
% biographies
%\newpage

%\begin{IEEEbiographynophoto}{Jane Doe}
%Biography text here.
%\end{IEEEbiographynophoto}

% You can push biographies down or up by placing
% a \vfill before or after them. The appropriate
% use of \vfill depends on what kind of text is
% on the last page and whether or not the columns
% are being equalized.

%\vfill

% Can be used to pull up biographies so that the bottom of the last one
% is flush with the other column.
%\enlargethispage{-5in}

\end{document}